---
title: "Kaggle 11.11.15"
output: html_document
---

####Prepping data
```{r}
setwd("/Users/Lowry/Documents/Stanford/Senior/Fall/Stats_202/Data_Directory")
training.features <- read.csv(file="training_features.csv", header=TRUE, sep=",")
training.target <- read.csv(file="training_target.csv",header=TRUE, sep=",")
feature.names <- names(training.features) # Create vector of feature names
num.col.nas <- 0
for (feature.name in feature.names[-1]) {
                    # Give the new dummy variable a meaningful name
                    dummy.name <- paste0("is.na.",feature.name)
                    is.na.feature <- is.na(training.features[,feature.name])
                    num.col.nas[feature.name] = sum(is.na(training.features[,feature.name]))
                    # Convert boolean values to binary
                    training.features[,dummy.name] <- as.integer(is.na.feature)
                    # Replace NA with median value for each column
                    training.features[is.na.feature,feature.name] <- median(training.features[,feature.name], na.rm = TRUE)
}
dim.data.frame(training.features) # Confirms dimension of new dataframe
merged.data=merge(training.target,training.features,by="subject.id") # creates one DF with ALSFRS_slope (response variable) as column 2
sum(num.col.nas)
```

Column 1: subject.id columns 2:858 original training.feature columns, with NAs replaced with median value columns 859-1715 new dummy columns named "is.na.[original column_name]"
```{r}
leaderboard.predictions <- read.csv(file = "leaderboard_predictions-example.csv", header=TRUE, sep=",")
leaderboard.features <- read.csv(file="leaderboard_features.csv", header=TRUE, sep=",")
feature.names <- names(leaderboard.features) # Create vector of feature names
num.col.nas.lead <- 0
for (feature.name in feature.names[-1]) {
                    # Give the new dummy variable a meaningful name
                    dummy.name <- paste0("is.na.",feature.name)
                    is.na.feature <- is.na(leaderboard.features[,feature.name])
                    # Convert boolean values to binary
                    leaderboard.features[,dummy.name] <- as.integer(is.na.feature)
                    # Replace NA with median value for each column
                    leaderboard.features[is.na.feature,feature.name] <- median(leaderboard.features[,feature.name], na.rm = TRUE)
}
dim.data.frame(leaderboard.features) # Confirms dimension of new dataframe
sum(num.col.nas.lead)
```

Removing variables from training and leaderboard data that have more than half of observations recorded as NA in training data set.
```{r}
training.df.cut <- merged.data[c("subject.id", "ALSFRS_slope")]
lead.df.cut <- leaderboard.features["subject.id"]
thrown.out <- c()
num.col.nas <- num.col.nas[2:length(num.col.nas)]
for(i in 1:length(feature.names)) {
  #if the count of the variable is greater than dim(training.features)[1]/2, discard
  if(num.col.nas[i] >= dim(training.features)[1]/2) {
    thrown.out <- c(thrown.out, names(num.col.nas[i]))
  } else {
    training.df.cut[,names(training.features[i])] <- training.features[,i]
    lead.df.cut[,names(training.features[i])] <- leaderboard.features[,names(training.features[i])]
  }
}
thrown.out
```

Removing variables from training and leaderboard data that have more than half of observations recorded as NA in training data set.
```{r}
remaining.features = names(lead.df.cut)
lead.col.num.nas = c()
lead.thrown.out = c()
for(remaining.feature in remaining.features) {
  lead.col.num.nas[remaining.feature] = sum(is.na(lead.df.cut[,remaining.feature]))
}

for(i in 1:length(remaining.features)) {
  #if the count of the variable is greater than dim(training.features)[1]/2, discard
  if(lead.col.num.nas[i] >= dim(lead.df.cut)[1]/2) {
    lead.thrown.out <- c(lead.thrown.out, names(lead.col.num.nas[i]))
    lead.df.cut = lead.df.cut[,-i]
    training.df.cut = training.df.cut[,-(i+1)]
  } 
}
lead.thrown.out
```

