---
title: "Calhoon_Kaggle_Submission_20151028-a"
author: "Ken Calhoon"
date: "October 16, 2015"
output: html_document
---

RMD shows code for Ridge and Lasso.

Code below loads feature and target datasets, adds dummy columns for NA values, adds the median variable value for NA values, and merges target and feature datasets. 

Column 1: subject.id
column 2: ALSFRS_slope
columns 3:859 original training.feature columns, with NAs replaced with median value
columns 860-1716 new dummy columns named "is.na.[original column_name]"

```{r, echo=FALSE}
setwd("~/Dropbox/Stats202/Kaggle")
training.features <- read.csv(file="training_features.csv", header=TRUE, sep=",")
training.target <- read.csv(file="training_target.csv",header=TRUE, sep=",")
feature.names <- names(training.features) # Create vector of feature names
for (feature.name in feature.names[-1]) {
                    # Give the new dummy variable a meaningful name
                    dummy.name <- paste0("is.na.",feature.name)
                    is.na.feature <- is.na(training.features[,feature.name])
                    # Convert boolean values to binary
                    training.features[,dummy.name] <- as.integer(is.na.feature)
                    # Replace NA with median value for each column
                    training.features[is.na.feature,feature.name] <- median(training.features[,feature.name], na.rm = TRUE)
}
dim.data.frame(training.features) # Confirms dimension of new dataframe
merged.data=merge(training.target,training.features,by="subject.id") # creates one DF with ALSFRS_slope (response variable) as column 2
```

Code below loads  the leaderboard_feature dataset, adds dummy columns for NA values, and adds the median variable value for NA values. Loads leaderboard.predictions

Column 1: subject.id
columns 2:858 original training.feature columns, with NAs replaced with median value
columns 859-1715 new dummy columns named "is.na.[original column_name]"

```{r, echo=FALSE}
setwd("~/Dropbox/Stats202/Kaggle")
leaderboard.predictions <- read.csv("~/Dropbox/Stats202/Kaggle/leaderboard_predictions-example.csv")
leaderboard.features <- read.csv(file="leaderboard_features.csv", header=TRUE, sep=",")
feature.names <- names(leaderboard.features) # Create vector of feature names
for (feature.name in feature.names[-1]) {
                    # Give the new dummy variable a meaningful name
                    dummy.name <- paste0("is.na.",feature.name)
                    is.na.feature <- is.na(leaderboard.features[,feature.name])
                    # Convert boolean values to binary
                    leaderboard.features[,dummy.name] <- as.integer(is.na.feature)
                    # Replace NA with median value for each column
                    leaderboard.features[is.na.feature,feature.name] <- median(leaderboard.features[,feature.name], na.rm = TRUE)
}
dim.data.frame(leaderboard.features) # Confirms dimension of new dataframe
```

Replicating Ridge code from book

```{r}
x=model.matrix(Salary ∼ .,Hitters)[,-1]
y=Hitters$Salary
library(glmnet)
grid=10^seq(10,-2,length=100)
ridge.mod=glmnet(x,y,alpha=0,lambda=grid)
```


Ridge code

```{r}
#install.packages("glmnet")
library(glmnet)
merged.data.subset=merged.data[,1:859]
x=model.matrix(ALSFRS_slope~.,merged.data.subset)[,-2]
y=merged.data.subset$ALSFRS_slope
grid=20^seq(10,-2,length=100)
ridge.mod=glmnet(x,y,alpha=0,lambda=grid,standardize=FALSE)
```

```{r}
dim(coef(ridge.mod))
ridge.mod$lambda[60]
coef(ridge.mod)[,60]
```

```{r}
set.seed(1)
train=sample(1:nrow(x),nrow(x)/2)
test=(-train) # shouldn't this be != instead?
y.test=y[test]
ridge.mod=glmnet(x[train,],y[train],alpha=0,lambda=grid,thresh=1e-12)
ridge.pred=predict(ridge.mod,s=4,newx=x[test,])
rmse=mean((ridge.pred-y.test)^2)^.5
rmse
```

Compare to lambda = 0 (means no influence from lambda)
```{r}
set.seed(1)
train=sample(1:nrow(x),nrow(x)/2)
test=(-train) # shouldn't this be != instead?
y.test=y[test]
ridge.mod=glmnet(x[train,],y[train],alpha=0,lambda=grid)
ridge.pred=predict(ridge.mod,s=0,newx=x[test,]) #change s value
rmse=mean((ridge.pred-y.test)^2)^.5
rmse
```


Lasso
```{r}
#install.packages("glmnet")
library(glmnet)
merged.data.subset=merged.data[1:1000,1:858] #add back all rows
x=model.matrix(ALSFRS_slope~.,merged.data.subset)[,-2]
y=merged.data.subset$ALSFRS_slope
lasso.mod=glmnet(x[train ,],y[train],alpha=1,lambda=grid)
plot(lasso.mod)
```

Lasso + CV--Test Code
```{r}

rmse
#rerun lasso.mod with all x & y (submission)
merged.data.subset=merged.data[,1:859] #add back all data
#predict with leaderboard.features (submission)

```

Lasso + CV--Real Model [Need to modify this]
```{r}
library(glmnet)
library(boot)
set.seed(1)
merged.data.subset=merged.data[1:100,1:858] #can reduce rows to make easier to work with
x=model.matrix(ALSFRS_slope~.,merged.data.subset)[,-2]
y=merged.data.subset$ALSFRS_slope
#ALSFRS_slope=y
train=sample(1:nrow(x),nrow(x)/2)
test=(-train)
y.test=y[test]
cv.out=cv.glmnet(x[train ,],y[train],alpha=1,nfolds=3) #change later to K=10, alpha can range 0 to 1 and can run CV on this to get best mix
plot(cv.out)
bestlam=cv.out$lambda.min
bestlam
#insert code to do the cv error estimate
lasso.mod=glmnet(x[train ,],y[train],alpha=1,lambda=bestlam)
cv.out=cv.glm(merged.data.subset[test,],lasso.mod,K=3) #Change this value later to 5 or 10
lasso.pred=predict(lasso.mod,s=bestlam ,newx=x)
rmse=mean((lasso.pred-y.test)^2)^.5
rmse

leaderboard.x=as.matrix(leaderboard.features[,1:858])
#leaderboard.x=model.matrix(~.,data=leaderboard.features[,1:858])
ALSFRS.leader.pred<-predict(lasso.mod,s=bestlam,newx=leaderboard.x) #make predictions

leaderboard.predictions$ALSFRS_slope <- ALSFRS.leader.pred
head(leaderboard.predictions)
```


Coefficients
```{r}
out=glmnet(x,y,alpha=1,lambda=grid)
lasso.coef=predict(out,type="coefficients",s=bestlam)
lasso.coef[lasso.coef!=0]
```

```{r}
out=glmnet(x,y,alpha=1,lambda=grid)
lasso.coef=predict(out,type="coefficients",s=bestlam)
lasso.coef[lasso.coef!=0]

lasso.pred=predict(lasso.coef,x,type="response")
rmse=mean((lasso.pred-y.test)^2)^.5
rmse
```

LARS: least angle regression



#####Predict ALSFRS_slope from leaderboard data and make predictions

Predict the ALSFRS_slope and load to leaderboard.predictions

```{r}
#rerun lasso.mod with all x & y (submission)
#predict with leaderboard.features (submission)
merged.data.subset=merged.data[,1:859] #add back all data

ALSFRS.leader.pred<-predict(lasso.mod,s=bestlam ,newx=leaderboard.features[,1:858]) #make predictions
leaderboard.predictions$ALSFRS_slope <- ALSFRS.leader.pred
head(leaderboard.predictions)
```

We use **write.csv** function to write a CSV file in the contest format with the leaderboard subject predictions. 

```{r}
write.csv(leaderboard.predictions, file = "leaderboard_predictions_20151023.csv",row.names=FALSE) # NEED TO MODIFY
```

