---
title: "Calhoon_Kaggle_Scratch_v1"
author: "Ken Calhoon"
date: "October 16, 2015"
output: html_document
---

Code below loads feature and target datasets, adds dummy columns for NA values, adds the median variable value for NA values, and merges target and feature datasets. 

Column 1: subject.id
column 2: ALSFRS_slope
columns 3:859 original training.feature columns, with NAs replaced with median value
columns 860-1716 new dummy columns named "is.na.[original column_name]"

```{r, echo=FALSE}
setwd("~/Dropbox/Stats202/Kaggle")
training.features <- read.csv(file="training_features.csv", header=TRUE, sep=",")
training.target <- read.csv(file="training_target.csv",header=TRUE, sep=",")
feature.names <- names(training.features) # Create vector of feature names
for (feature.name in feature.names[-1]) {
                    # Give the new dummy variable a meaningful name
                    dummy.name <- paste0("is.na.",feature.name)
                    is.na.feature <- is.na(training.features[,feature.name])
                    # Convert boolean values to binary
                    training.features[,dummy.name] <- as.integer(is.na.feature)
                    # Replace NA with median value for each column
                    training.features[is.na.feature,feature.name] <- median(training.features[,feature.name], na.rm = TRUE)
}
dim.data.frame(training.features) # Confirms dimension of new dataframe
merged.data=merge(training.target,training.features,by="subject.id") # creates one DF with ALSFRS_slope (response variable) as column 2
```

Code below loads  the leaderboard_feature dataset, adds dummy columns for NA values, and adds the median variable value for NA values. Loads leaderboard.predictions

Column 1: subject.id
columns 2:858 original training.feature columns, with NAs replaced with median value
columns 859-1715 new dummy columns named "is.na.[original column_name]"

```{r, echo=FALSE}
setwd("~/Dropbox/Stats202/Kaggle")
leaderboard.predictions <- read.csv("~/Dropbox/Stats202/Kaggle/leaderboard_predictions-example.csv")
leaderboard.features <- read.csv(file="leaderboard_features.csv", header=TRUE, sep=",")
feature.names <- names(leaderboard.features) # Create vector of feature names
for (feature.name in feature.names[-1]) {
                    # Give the new dummy variable a meaningful name
                    dummy.name <- paste0("is.na.",feature.name)
                    is.na.feature <- is.na(leaderboard.features[,feature.name])
                    # Convert boolean values to binary
                    leaderboard.features[,dummy.name] <- as.integer(is.na.feature)
                    # Replace NA with median value for each column
                    leaderboard.features[is.na.feature,feature.name] <- median(leaderboard.features[,feature.name], na.rm = TRUE)
}
dim.data.frame(leaderboard.features) # Confirms dimension of new dataframe
```


####Initial Models

Top 12 correlation variables
```{r, echo=FALSE}
library(boot)
set.seed(1)
glm.fit=glm(ALSFRS_slope~onset_delta+last.q1_speech+fvc.slope+min.q1_speech+meansquares.q1_speech+sd.alsfrs_total+last.mouth+min.mouth+mean.q1_speech+meansquares.mouth+onset_site.Limb+mean.mouth, data = merged.data)
cv.err=cv.glm(merged.data,glm.fit, K=5) #5 cross-folds
# cv.err$delta # cross validation sqd error estimate (the second value adjusts for not using LOOCV)
cv.err$delta^.5 #cv estimate of RMSE
ALSFRS.pred<-predict(glm.fit,merged.data,type="response") #make predictions
plot(ALSFRS.pred,merged.data$ALSFRS_slope) #plot the results
RMSE=(mean((ALSFRS.pred-merged.data$ALSFRS_slope)^2))^.5
RMSE
```


~ Best subset model from 12 variables above. 4 variables
```{r, echo=FALSE}
library(boot)
set.seed(1)
glm.fit=glm(ALSFRS_slope~onset_delta+last.q1_speech+fvc.slope+sd.alsfrs_total, data = training.subset.with.target)
cv.err=cv.glm(training.subset.with.target,glm.fit, K=5) #5 cross-folds
# cv.err$delta # cross validation sqd error estimate (the second value adjusts for not using LOOCV)
cv.err$delta^.5 #cv estimate of RMSE
ALSFRS.pred<-predict(glm.fit,training.subset.with.target,type="response") #make predictions
plot(ALSFRS.pred,training.subset.with.target$ALSFRS_slope) #plot the results
plot(ALSFRS.pred,training.subset.with.target$ALSFRS_slope-ALSFRS.pred) # plot the residuals
RMSE=(mean((ALSFRS.pred-training.subset.with.target$ALSFRS_slope)^2))^.5
RMSE
summary(glm.fit)
#should rerun the model with all data
```

(new) RMSE calc on the validation set
```{r}
valid.pred<-predict(glm.fit,validation.subset.with.target,type="response")
valid.rmse=mean((valid.pred-validation.target$ALSFRS_slope)^2)^.5
valid.rmse
```

#####Notes
Reran this model.
CV error=.578
Training RMSE=0.577
Validation Set RMSE=.612
Kaggle=.5882


#####Predict ALSFRS_slope from leaderboard data and make predictions

Predict the ALSFRS_slope and load to leaderboard.predictions

```{r}
ALSFRS.leader.pred<-predict(glm.fit,leaderboard.features,type="response") #make predictions
leaderboard.predictions$ALSFRS_slope <- ALSFRS.leader.pred
head(leaderboard.predictions)
```

We use **write.csv** function to write a CSV file in the contest format with the leaderboard subject predictions. 

```{r}
write.csv(leaderboard.predictions, file = "leaderboard_predictions_20151023.csv",row.names=FALSE) # NEED TO MODIFY
```

